
# Personal Activity Monitoring
#### Girish Doshi
#### Friday, April 15, 2015

### Loading and preprocessing data 
```{r}
raw_data <- read.csv("repdata-data-activity/activity.csv")

library(dplyr)
no_na_data <- raw_data[complete.cases(raw_data), ]

```

### What is mean total number of steps  per day?

1 Calculate the total number of steps taken per day
```{r}
steps_per_day <- group_by(no_na_data, date) %>% summarise(Total.steps = sum(steps))
```

2 Histogram of total number of steps taken each day
```{r}
hist(steps_per_day$Total.steps, xlab = "Total steps per day", main = "Histogram of Total number of steps each day")

barplot(steps_per_day$Total.steps, main="Total Steps taken per day", names.arg=steps_per_day$date, width=5, xlab = "date", ylab = "No. of steps")
```

3 The mean and median of the total number of steps taken per day
```{r}
summary(steps_per_day)
```

### What is the average daily activity pattern?
1 Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
avg_steps_per_interval_accross_all_days <- group_by(no_na_data, interval) %>% 
                                                summarise(avg.steps = mean(steps))

plot(avg_steps_per_interval_accross_all_days$interval,
     avg_steps_per_interval_accross_all_days$avg.steps,
     type="l",
     xlab="5-minite interval",
     ylab="Average no. of steps",
     main= "Average steps taken per 5-minute interval accross all days")

```

2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
filter(avg_steps_per_interval_accross_all_days, avg.steps == max(avg.steps))
```

### Imputing missing values

1 Total number of missing values in the 'steps' column in the raw data set
```{r}
sum(is.na(raw_data$steps))

```

2 Fill the missing values  
   We fill missing values with average steps per interval accross all days.

```{r}
filled_df <- mutate(raw_data, avg.steps = rep(avg_steps_per_interval_accross_all_days$avg.steps, (nrow(raw_data)/length(unique(raw_data$interval)))))  %>% 
        mutate(steps = ifelse(is.na(steps), avg.steps, steps)) %>% 
        select(1:3)

```

3 Histogram of total number of steps taken each day

```{r}
filled_steps_per_day <- group_by(filled_df, date) %>% summarise(Total.steps = sum(steps))

hist(filled_steps_per_day$Total.steps, xlab = "Total steps per day", main = "Histogram of Total number of steps each day")

barplot(filled_steps_per_day$Total.steps, main="NA Values filled \n Total Steps taken per day", names.arg=filled_steps_per_day$date, width=5, xlab = "date", ylab = "No. of steps")

```

4 The mean and median of the total number of steps taken per day
```{r}
summary(steps_per_day)
```

### Are there differences in activity patterns between weekdays and weekends?

```{r}
data_by_days <- mutate(filled_df, day = factor( weekdays(as.Date(date)) %in% c("Saturday", "Sunday"), labels = c("weekend", "weekday"))) %>% 
        group_by(day, interval) %>%
        summarise(avg.steps = mean(steps))

library(lattice)
xyplot(avg.steps ~ interval | day, data_by_days, layout = c(1, 2), type = "l")
```
